---
title: "Quiz: Intro to R"
author: Ifeoma Egbogah
output: learnr::tutorial
runtime: shiny_prerendered
tutorial:
  id: week1-quiz
  description: "Covers RStudio, basic syntax, data types, and data import"
---

```{r setup, include=FALSE}

library(learnr)
library(gradethis)

gradethis::gradethis_setup()
tutorial_options(exercise.checker = gradethis::grade_learnr)

knitr::opts_chunk$set(echo = TRUE)

```


## Welcome to Week 1!

In this tutorial, you'll:

* Install R/RStudio

* Learn basic syntax

* Work with R data types and structures

* Practice importing data

---

## Install R

### Test your knowledge
```{r quiz1, echo=FALSE}

quiz(caption = "Quiz - Understanding R and RStudio",
  question("Is R free to download and use?",
    answer("Yes!", correct = TRUE, message = "R is free _and_ open source, which means that anyone can read, use, and copy the source code of the R language."),
    answer("No.")
  ),
  question("Where can you download the latest version of R?",
    answer("https://cran.r-project.org", correct = TRUE),
    answer("https://tidyverse.org", correct = FALSE),
    answer("https://posit.co/rstudio", correct = FALSE),
    answer("https://github.com", correct = FALSE),
    allow_retry = TRUE
  ),
  question("How often should you update R?",
    answer("Everytime you use it", message = "This will be too often unless you use R very rarely!"),
    answer("About once a year", correct = TRUE, "A new version of R is released about once a year. Update sooner if you encounter a bug that you cannot explain."),
    answer("Never", message = "A new version of R is released about once a year. I'll assume that you are using the newest version of R, which will be the fastest version with the fewest unexpected behaviors." ),
    allow_retry = TRUE
  )
)

```

## Install RStudio

### Test your knowledge

```{r quiz2, echo=FALSE}

quiz(caption = "Quiz - Install RStudio",
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use R.", correct = TRUE, message = "RStudio organizes your workspace and makes it easier to write, use, debug, and save R code. I highly recommend it and will rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code", message = "And thank goodness! Code provides a reproducible record of your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel."),
    answer("Another name for R", message = "R and RStudio are two separate things. R is a language, like English. Think of RStudio as a program that helps you use the language, kind of like how a word processing program helps you write in English."),
    allow_retry = TRUE
  ),
  question("Is the RStudio IDE free to download and use?",
    answer("Yes!", correct = TRUE, message = "Like R, RStudio is free and open-source. There do exist professional versions of RStudio with enhanced features, but we will not rely on those features in these tutorials."),
    answer("No.", message = "Like R, RStudio is free and open-source. There do exist professional versions of RStudio with enhanced features, but we will not rely on those features in these tutorials.")
  ),
  question("Where do you download RStudio?",
    answer("www.rstudio.com/download", correct = TRUE, message = "For these tutorials, download and install the RStudio Desktop Open Source License."),
    answer("[cloud.r-project.org](http://cloud.r-project.org)", message = "This is where you download R, not RStudio."),
    answer("www.r-project.org"),
    answer("[cran.rstudio.org](http://cran.rstudio.org)"),
    allow_retry = TRUE
  ),
  question("Do you need to install R if you already have RStudio?",
    answer("Yes.", correct = TRUE),
    answer("No.", message = "R does not come with RStudio; you need to install R separately.")
  )
)
```

## Install Packages

### Test your knowledge

```{r names, echo = FALSE}
quiz(caption = "Quiz - Working with Packages",
  question("What command do you use to install packages?",
    answer("`library()`", message = "We will see what `library()` does later."),
    answer("`install.packages()`", correct = TRUE),
    answer("`install_packages()`"),
    answer("There is no command. You must visit [cran.r-project.org](http://cran.r-project.org) and download packages manually.", message = "R makes it easy to download packages. You connect to the internet and then run one of the commands above."),
    allow_retry = TRUE
  ),
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R"),
    answer("Every time you restart your computer"),
    answer("Only once. Afterwards, R can find it on your hard drive as needed.", correct = TRUE),
    answer("Never, as long as you are connected to the internet.", message = "This could be true if you are using R over a cloud service. However, if you are using R locally on your own computer, you will need to install each package that you use on your computer."),
    allow_retry = TRUE
  ),
  question("What is the tidyverse?",
    answer("A collection of packages that work well together and provide tools for common data science tasks.", correct = TRUE, message = 'You can install all of the tidyverse packages at once with `install.packages("tidyverse")`.'),
    answer("A word processing program for R"),
    answer("A code style that reads like poetry"),
    answer("An alternate universe where everyone is an R programmer."),
    allow_retry = TRUE
  )
)
```


## Data Type Recognition

```{r datatype-question, echo=FALSE}
quiz(
  question("Which of the following is a logical value in R?",
           answer("TRUE", correct = TRUE),
           answer("42L", correct = FALSE),
           answer("\"Data\"", correct = FALSE),
           allow_retry = TRUE),
  question("Which of the following is not a basic data types in R?",
    answer("Character"),
    answer("Integer"),
    answer("Logical"),
    answer("Data Frame", correct = TRUE, message = "Data frames are data structures, not atomic types."),
    allow_retry = TRUE)
)
```


### 
```{r quiz3, exercise=TRUE}

# Assign your name to an object called your_name
# Then use typeof() to check what data type it is
```

```{r quiz3-solution, echo=FALSE}
grade_code({
  pass_if(
    exists("your_name") &&
    is.character(your_name),
    "✅ Looks good!"
  )
  fail("Did you assign your name to a variable using quotes and check its type?")
})

```


## Code Exercise: Basic Arithmetic

```{r basic-addition, exercise=TRUE}

# Add 10 and 15 and assign it to a variable called `result`
# Then print the result

```

```{r basic-addition-solution, echo=FALSE}

grade_code(
  correct = ~ identical(result, 25),
  incorrect = ~ "Try again. Did you correctly add 10 and 15 and assign it to `result`?"
)

```


```{r math-op-question, exercise = TRUE}

a <- 3
b <- 2
c <- a^b

```



```{r math-op, echo=FALSE}

quiz(
  question("What value is printed when the code above is run?",
    answer("6", correct = FALSE),
    answer("9", correct = TRUE),
    answer("5", correct = FALSE),
    answer("1.5", correct = FALSE),
    allow_retry = TRUE
  )
)


```

## Data Structure

```{r vector-creation, exercise=TRUE}
# Create a numeric vector with values 1, 2, 3, 4

```


```{r vector-creation-solution, echo=FALSE}
grade_code(
  correct = ~ my_vector <- c(1, 2, 3, 4),
  incorrect = ~ "Try again"
)
```

```{r ds_identify, echo=FALSE}

quiz(
  question("Which of the following is not a data structure in R?",
    answer("Vector"),
    answer("Matrix"),
    answer("Factor", correct = TRUE),
    answer("List"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question("A vector in R can contain elements of different types.",
    answer("TRUE"),
    answer("FALSE", correct = TRUE),
    allow_retry = TRUE
  ),
  question("What is a key characteristic of a matrix in R?",
    answer("Each column can be a different data type"),
    answer("All elements must be the same type", correct = TRUE),
    answer("It must have row names"),
    answer("You can only create it with the data.frame() function"),
    allow_retry = TRUE)
)
```


```{r ds_factors, exercise=TRUE}
# Create a factor variable called colours with values: "red", "blue", "red", "green"
# Then print the levels of the factor

```

```{r ds_factors-solution, echo=FALSE}

colours <- factor(c("red", "blue", "red", "green"))
levels(colours)

```



```{r ds_factors-check, echo=FALSE}

grade_code({
  # check if 'colours' exists
  if (!exists("colours")) {
    fail("You must create a variable named 'colours'.")
  }

  # check if it's a factor
  if (!is.factor(colours)) {
    fail("'colours' must be a factor.")
  }

  # check the levels — order doesn’t matter
  expected_levels <- c("red", "blue", "green")
  actual_levels <- levels(colours)

  if (!all(expected_levels %in% actual_levels)) {
    fail("Your factor is missing one or more expected levels.")
  }

  pass("✅ Great! You created a factor and included the correct levels.")
})

```


```{r ds_df_create, exercise=TRUE}

# Create a data frame called books with the following:
# title: c("A", "B")
# year: c(2001, 2005)
# available: c(TRUE, FALSE)

```


```{r ds_df_create-solution, echo=FALSE}

grade_code(
  correct = ~ books <- tibble(title = c("A", "B"),
                                  year= c(2001, 2005),
                                  available = c(TRUE, FALSE)),
  incorrect = ~ "Try again"
)

```


## Challenge: Read Data from CSV

```{r import-csv, exercise=TRUE}
# Load the readr package and import a CSV from URL

# https://pos.it/r4ds-students-csv

```


```{r import-csv-solution, echo=FALSE}

grade_code(
  correct = ~ students <- read_csv("https://pos.it/r4ds-students-csv", na = c("N/A", "")),
  incorrect = ~ "Try again"
)

students
```

### Importing Data from CSV and Excel
You have a file `data/employees.csv` with column names in row 3 (the first two rows are metadata). Which `readr` function call reads the data correctly?

```{r import, echo=FALSE}

quiz(
  question("How do you import a CSV with metadata in the first two rows?",
    answer("read_csv('data/employees.csv')", correct = FALSE),
    answer("read_csv('data/employees.csv', skip = 2)", correct = TRUE),
    answer("read_csv('data/employees.csv', col_names = FALSE)", correct = FALSE),
    answer("read_csv('data/employees.csv', na = 'NA')", correct = FALSE),
    allow_retry = TRUE
  )
)
```

