---
title: "Week 2: Data Cleaning and Wrangling"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

library(learnr)
library(tidyverse)
library(here)

tutorial_options(exercise.timelimit = 60)
```

# Week 2

Welcome to **Week 2** of your data science journey! In this module, you'll learn how to clean and reshape your data using `tidyverse` tools.

> 💡 Remember: Most datasets in the real world are *messy*. Your job is to tidy them for effective analysis.



### What is Tidy Data?

Tidy data has:
- Each **variable** in a column
- Each **observation** in a row
- Each **value** in a single cell

::: {.alert alert-info}
📌 This standard makes your data easier to manipulate, visualize, and model.
:::

## Reshaping Data

### Columns with Multiple Values

Use `separate()` 

```{r tidy-question, exercise=TRUE}
# Fix this dataset to be tidy.
# Try splitting Name_Age and Fever_Symptom into two separate columns
# Hint: Use separate() to fix the duration column

patient_summary <- tibble::tibble(
  Name_Age = c("John-32", "Lara-45", "Ben-28", "Zoe-40"),
  Gender = c("M", "F", "M", "F"),
  Visit_Date = c("2025-08-01", "2025-08-01", "2025-08-02", "2025-08-02"),
  Fever_Symptom = c("Yes-High", "No-", "Yes-Low", "Yes-Medium")
)


# Your code here
```


```{r tidy-question-solution, echo=FALSE}
patient_summary %>%
  separate(Name_Age, into = c("Name", "Age"), sep = "-") %>%
  separate(Fever_Symptom, into = c("Fever", "Symptom"), sep = "-")
```

```{r reshape-question, exercise=TRUE}
# Separate time into hour and minute

flight_schedule <- tibble::tibble(
  flight = c("AZ101", "BA202", "DL303"),
  departure = c("08:45 AM", "12:15 PM", "06:30 PM"),
  gate = c("A1", "B3", "C2")
)

# Your code here
```


```{r reshape-question-solution, echo=FALSE}
flight_schedule %>%
  separate(departure, into = c("hour", "minute"), sep = ":")
```


```{r reshape-question2, exercise=TRUE}

# Separate the minute into minutes and time of day

flight_schedule <- tibble::tibble(
  flight = c("AZ101", "BA202", "DL303"),
  departure = c("08:45 AM", "12:15 PM", "06:30 PM"),
  gate = c("A1", "B3", "C2")
)%>%
  separate(departure, into = c("hour", "minute"), sep = ":")

# Your code here

```

```{r reshape-question2-solution, echo=FALSE}

flight_schedule %>%
  separate(minute, into = c("minute", "time_of_day"), sep = " ")

```

### Separate Rows from Multi-Value Cell
Use `separate_rows()` when multiple values in a cell should each be in a row.
```{r multi-Value, exercise=TRUE}
# Separate student names into rows

school_subjects <- tibble::tibble(
  subject = c("Math", "Science", "Literature"),
  students = c("Ali, Bola, Chi", "Tola, Emma", "Ngozi, Ken, Lara")
)

# Your code here

```


```{r multi-Value-solution, echo=FALSE}
school_subjects %>%
  separate_rows(students, sep = ", ")
```

## Combining Columns

Use `unite()` to merge columns into one.

```{r combine-question, exercise=TRUE}
# Combine first and last name into a 'name' column separated by a space

customer_name <- tibble::tibble(
  first_name = c("Mary", "James", "Aisha"),
  last_name = c("Johnson", "Brown", "Ahmed"),
  country = c("USA", "UK", "Nigeria")
)


# Your code here
```


```{r combine-question-solution, echo=FALSE}
customer_name %>%
  unite("full_name", first_name, last_name, sep = " ")
```



You work for a multinational company that uses auto-dialer software to contact its customers. When new customers subscribe online, they are asked for a phone number but they often forget to add the country code needed for international calls. You were asked to fix this issue in the database. You’ve been given a data frame with national numbers and country codes named phone_nr_df. Now you want to combine the country_code and national_number columns to create valid international numbers.

```{r international-numbers, exercise=TRUE}

phone_nr_df <- tribble(~country, ~country_code, ~national_number,
                       "USA",            +1, 2025550117,     
                       "United Kingdom", +44, 1632960924,    
                       "Brazil",         +55, 95552452220, 
                       "Australia",      +61, 1900654321, 
                       "China",          +86, 13555953217, 
                       "India",          +91, 8555843898)

# Your code here

```

```{r international-numbers-solution, echo=FALSE}


phone_nr_df %>% 
  unite("international_number", country_code, national_number, sep = "")
```

The drink_df data set includes a variable that has the quantities and units of ingredients used in making each drink. Now you want to create an overview of how much of each ingredient you should buy to make these drinks.

Inspect drink_df and find the right separator in the ingredient's column.
Separate the ingredients column so that for each drink and each ingredient gets a row.

```{r drink, exercise=TRUE}

drink_df <- tribble(~drink, ~ingredients,                                 
                    "Chocolate milk",  "milk 0.3 L; chocolate 40 g; sugar 10 g",  
                    "Orange juice",   "oranges 3; sugar 20 g",                  
                    "Cappuccino",     "milk 0.1 L; water 0.1 L; coffee 30 g; sugar 5 g")


# Your code here
```


```{r drink-solution, echo=FALSE}

drink_df %>% 
  separate_rows(ingredients, sep = "; ", convert = TRUE)

```

- Inspect the output of the previous step to find the separator that splits the ingredients column into three columns: ingredient, quantity, and unit.
- Make sure to convert data types to numeric when possible.
- Group the data by ingredient and unit.
- Calculate the total quantity of each ingredient.


```{r drink2, exercise=TRUE}

drink_df <- tribble(~drink, ~ingredients,                                 
                    "Chocolate milk",  "milk 0.3 L; chocolate 40 g; sugar 10 g",  
                    "Orange juice",   "oranges 3; sugar 20 g",                  
                    "Cappuccino",     "milk 0.1 L; water 0.1 L; coffee 30 g; sugar 5 g")

# Your code here

```



```{r drink2-solution, echo=FALSE}

drink_df %>% 
  separate_rows(ingredients, sep = "; ") %>% 
  separate(ingredients, into = c("ingredients", "quantity", "unit"), sep = " ", convert = TRUE) %>% 
  group_by(ingredients, unit) %>% 
  summarise(total = sum(quantity))

```


## Missing Values

Use `is.na()`, `sum()`, `complete.cases()`, `replace_na()`, and `fill()` to identify and handle missing data.

```{r missing-values, exercise=TRUE}
# Count NAs

covid_cases <- tibble::tibble(
  state = c("Lagos", "Kano", "Abuja", "Enugu"),
  confirmed = c(2345, NA, 1500, NA),
  recovered = c(1200, 800, NA, NA)
)


# Your code here
```


```{r missing-values-solution, echo=FALSE}
sum(is.na(covid_cases))
```


### Impute with zeros
```{r missing-values2, exercise=TRUE}

# Replace NA with 0

covid_cases <- tibble::tibble(
  state = c("Lagos", "Kano", "Abuja", "Enugu"),
  confirmed = c(2345, NA, 1500, NA),
  recovered = c(1200, 800, NA, NA)
)

# Your code here

```



```{r missing-values2-solution, echo=FALSE}
covid_cases %>%
  replace_na(list(confirmed = 0, recovered = 0))
```


### 'fill'

You’ve been asked to create a report that allows management to compare sales figures per quarter for two years. The problem is that the dataset **(sales_df)** contains missing values. You’ll need to impute the values in the year column so that you can visualize the data.

- Inspect sales_df in the console, pay attention to the year column.
- Use the `fill()` function to impute the year column in the correct direction.
- Create a line plot where each year has a different color.


```{r fill, exercise=TRUE}
sales_df<- tribble(~year, ~quarter, ~sales,
                    NA, "Q1", 12498,
                   NA, "Q2", 20461,
                   NA, "Q3", 19737,
                   2019, "Q4", 20314,
                   NA, "Q1", 13494,
                   NA, "Q2", 19314,
                   NA, "Q3", 23640,
                   2020, "Q4", 22920)

# Your code here
```

```{r fill-solution, echo=FALSE}

sales_df %>%
  fill(year, .direction = "up") %>%    # Impute the year column
  ggplot(aes(x = quarter, y = sales, color = year, group = year)) +
  geom_line() # Create a line plot with sales per quarter colored by year.


```

## Pivoting: From Wide to Long

Use `pivot_longer()` when column names represent variables (e.g. countries).

```{r pivot-question, exercise=TRUE}
# Convert to long format

sales_wide <- tibble::tibble(
  region = c("North", "South", "East"),
  Q1 = c(10000, 12000, 9500),
  Q2 = c(11000, 12500, 10500),
  Q3 = c(11500, 13000, 10200),
  Q4 = c(12000, 13500, 11000)
)


# Your code here
```


```{r pivot-question-solution, echo=FALSE}

sales_wide %>%
  pivot_longer(cols = Q1:Q4, names_to = "quarter", values_to = "sales")

```



## Integrate Data with Joins
### Perform a left join

```{r left-join, exercise=TRUE}

# Perform a left join

employee_info <- tibble::tibble(
  emp_id = c(101, 102, 103),
  name = c("Tunde", "Ada", "Ola"),
  department = c("Logistics", "HR", "Finance")
)

employee_salary <- tibble::tibble(
  emp_id = c(101, 102, 104),
  monthly_salary = c(250000, 200000, 300000)
)

# Your code here
```



```{r left-join-solution, echo=FALSE}
employee_info %>%
  left_join(employee_salary, by = "emp_id")
```


## Challenge

### 

```{r quiz2, echo=FALSE}

quiz(
  question("Which of these is NOT part of tidy data?",
    answer("Each variable is in its own column"),
    answer("Each observation is in a column", correct = TRUE),
    answer("Each value is in its own cell"),
    answer("Each row is an observation")
  ),
  question("Which function is used to split a column?",
    answer("unite()"),
    answer("separate()", correct = TRUE),
    answer("pivot_wider()"),
    answer("fill()")
  )
)
```

## Summary

Today you learned how to:
- Structure your data using `tidyverse`
- Use `separate()`, `separate_rows()`, `unite()`
- Handle missing values with `replace_na()` and `fill()`
- Reshape your data using `pivot_longer()`

You’re one step closer to becoming a tidyverse pro! 🧹

Try applying these skills to your own dataset using `read_csv()` and wrangle it based on what you’ve learned!


